Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices

### 2022-06-30 paper.tex
[add] initial version of DoS section

%-------------------------------------------------------------------------- sec
\section{\uppercase{Denial of Service}}
\label{sec:dos}
\mEdhoc{} includes no specific measures for countering Denial of Service (DoS)
attacks and delegates the responsibility to handle this to the application.
%
\mEdhoc{} itself has limited functionality, few negotiable parameters and
processing rules
%
Hence, we argue in this section that the attack surface
is small from a DoS perspective.
%
However, \mEdhoc{} leaves much of the semantics of its messages undefined.
%
This increases the risk of applications vulnerable to DoS even if \mEdhoc{} is
arguably not vulnerable to DoS itself.
%

Section 8.7 of the \mEdhoc{} specification explains the design philosophy w.r.t.
DoS in two paragraphs and gives two recommendations.
%
The first being that mitigation of DoS based on sending proper messages should
be handled by lower layers, e.g., by checking reachability of the sending
party\footnote{Karl: this may be another way to perform reflection DoS attacks}.
%
The second being that an application may try to determine (in some way) that an
apparently valid message is in fact probably a forgery and should be ignored or
cause processing to stop.
%

>>>>

%-------------------------------------------------------------------------- sec
\section{\uppercase{Denial of Service}}
\label{sec:dos}
\mPoint{DoS is out of scope for \mEdhoc{}; apps to deal with it}
\mEdhoc{} includes no specific measures for countering Denial of Service (DoS)
attacks and delegates the responsibility to handle this to the application.
%
\mEdhoc{} itself has limited functionality, few negotiable parameters and
processing rules
%
Hence, we argue in this section that the attack surface
is small from a DoS perspective.
%
However, \mEdhoc{} leaves much of the semantics of its messages undefined.
%
This increases the risk of applications vulnerable to DoS even if \mEdhoc{} is
arguably not vulnerable to DoS itself.
%

\mPoint{\mEdhoc{} gives a little info of questionable use for developers}
Section 8.7 of the \mEdhoc{} specification explains the design philosophy w.r.t.
DoS in two paragraphs and gives two recommendations.
%
The first being that mitigation of DoS based on sending proper messages should
be handled by lower layers, e.g., by checking reachability of the sending
party\footnote{Karl: this may be another way to perform reflection DoS attacks}.
%
The second being that an application may try to determine (in some way) that an
apparently valid message is in fact probably a forgery and should be ignored or
cause processing to stop.
%

\mPoint{Responder is more vulnerable than initiator}.
For the same reasons as given in the IKEv2 DDoS analysis
by Nir and Smyslov~\cite{rfc8019} the responder is more vulnerable than the
initiator.
%
From a roles and message passing perspective, IKEv2 is very similar to all
methods in \mEdhoc{}.
%
The RFC provides much good advice for protocols with this structure that
\mEdhoc{} using applications can reuse.
%
We should look more into this.
%

\mPoint{Amplification}
An adversary can fill a message 1 with easily generated bogus data and send this
to a victim responder.
%
There is little work for the initiator, but both symmetric and asymmetric key
operations required by the responder.
%
\mEdhoc{} does not protect against this, but refer to the application and its
transport layer to resolve the problem.
%
This is a design decision, see Section 8.7 of \mEdhoc{} specification.
%
In addition to the computational resources the responder spends, it may also
need to obtain certificate chains and verify certificate revocation servers as a
result of the spoofed message.
%
The asymmetry between the adversaries effort and the responder may hence be
significant.
%
Since \mEdhoc{} specifically targets constrained devices this asymmetry needs to
be considered even more carefully than for IKEv2~\cite{rfc8019}.
%

Recommend that the role of responder is chosen carefully.
%
The responder is most vulnerable.
%
For example, is it better that a lock or one of the key cards takes on the
responder role and becomes vulnerable?
%

\mPoint{Reflection Attacks}
Reflection attacks are attacks where an adversary spoofs the source address of a
victim in a message and thereby tricks the responder to send a message to the
victim.
%
\mEdhoc{} leaves protection against this to the transport protocol and
application, and suggests using a path reachability method similar to IKEv2.
%
In situations where the responder does not need to perform additional
communications to verify certificates or similar, a reachability mechanism may
be more expensive than performing the computations.
\footnote{Find ref of that computation is often cheaper than radio
transmission, at least energy-wise.}
%
However, if round-trip times are large, and half-open states hence are kept
by the receiver for a long time, a reachability mechanism can trade
communication for storage.
\footnote{RFC8019 discuss some numbers as examples of tradeoffs. Should we set
up a simple model and derive an expression for trade-offs? Could be useful for
implementors to consider, but low academic value. This would capture also
amplification.}
%

\mPoint{In a constrained device randomness may be easily depleted}
Assume an adversary generate (deterministic) bogus message 1s and send these at
a sufficiently high rate to the responder.
%
\mEdhoc{} specifically targets constrained devices, that presumably may not
always have good randomness sources.
%
Is it possible to deplete or significantly reduce the entropy of the source and
in this way be able to predict the responder's random input to other sessions
(with other initiators)?
%
The LAKE WG have been talking about effects of poor randomness on signatures, so
this should be an issue, no?
%

\mPoint{\mEdhoc{} can deadlock if not used correctly}
An adversary may send forged error messages, and specifically reject proposed
ciphersuites.
%
The initiator is recommended to not try that ciphersuite with the responder
again in a new session for a while.
%
If the adversary rejects all options for a initiator this way, it can prevent
communication between the initiator and that responder until the cache is
cleared.
%
Potentially, an adversary could lock a party of a group out of communication all
together this way.
%
This needs to be considered by applications.
%

\mPoint{Is any of the methods better than the others from a DoS perspective?}
We considered this for quite some time, but there seems to be little difference
in this regard.
%
